### 2022-09-02 main/chapter_searches.tex
add some history of frequentism text

Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
proportion of stuff within a larger whole, which when formalized as a ratio
of measures and has broad applications.
Frequentist theory confines itself to the special case where those measures
are numbers of discrete randomly occurring events, such that a frequentist
probability is a ratio of frequencies:
\begin{equation}
\label{eqn:searches_prob_frequency_ratio}
\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
\end{equation}
where $n$ is the number observed to satisfy conditions $a$ and $b$ when
$N$ have satisfied $b$.
Interpreted with the Binomial distribution of
Equation~\ref{eqn:searches_binomial}, this definition converges to our more
general probability whenever it applies.

@@

Tom Stoppard's play `Rosencrantz and Guildenstern Are Dead'~\cite{
stoppard1967rosencrantz
}
contains a clearer example ---
Rosencrantz and Guildenstern toss a coin ninety-two times, all landing heads.
They play a game with a third player, landing eight more heads.
Is the next toss heads or tails?
The answer is not random --- it is burned in print onto paper, so no frequency
can be assigned.
Were subjective probabilities permitted, you could predict the result, and that
prediction would be subjectively informed by it being the one-hundred-and-first
toss, our cultural entanglement with in base-10 numbers, and your reading of
these words.
Frequentist theory forbids probabilities of such non-random unknowns, and is
left seeking alternatives.

Venn~\cite{venn1866logic},
Boole~\cite{boole1854investigation},
Kolmogorov~\cite{
kolomogoroff1933de,
kolomogoroff1950translated
},

@@

\TODO{
Fisher:
criteria for estimators,
(consistency, efficiency, sufficiency)
~\cite{fisher1922estimators}
testing, p-value,
likelihood, also ratios of p-values (here named probable errors)
(introduced the term likelihood? cite?)
~\cite{fisher1921probable}
tail area = significance
~\cite{fisher1925smrw}
}
Figure~\ref{fig:searches_fisher_likelihood}

>>>>

Probability as defined in Equation~\ref{eqn:searches_prob_ratio} is a
proportion of stuff within a larger whole, which when formalized as a ratio
of measures and has broad applications.
Frequentist theory confines itself to the special case where those measures
are numbers of discrete randomly occurring events, such that a frequentist
probability is a ratio of frequencies:
\begin{equation}
\label{eqn:searches_prob_frequency_ratio}
\freq{b}{a} = \lim_{N \to \infty}\frac{n}{N},
\end{equation}
where $n$ is the number observed to satisfy conditions $a$ and $b$ when
$N$ have satisfied $b$.
Interpreted with the Binomial model of
Equation~\ref{eqn:searches_binomial}, this definition converges to our more
general probability whenever it applies.

@@

Tom Stoppard's play `Rosencrantz and Guildenstern Are Dead'~\cite{
stoppard1967rosencrantz
}
contains a clearer example ---
Rosencrantz and Guildenstern toss a coin ninety-two times, all landing heads
(by ``un-, sub- or supernatural forces'').
They play a game with a third player, landing eight more heads.
Is the next toss heads or tails?
The answer is not random --- it is burned in print onto paper, so no frequency
can be assigned.
Were subjective probabilities permitted, you could predict the result, and that
prediction would be subjectively informed by it being the one-hundred-and-first
toss, our cultural entanglement with round numbers in base-10, and your reading
of these words.
Frequentist theory, however, forbids such subjective probabilities of
non-random unknowns, so is left seeking alternatives.

As established in Section~\ref{sec:searches_data_analysis}, the original
probability theory of Laplace and Bayes, did not require frequency
interpretations; these were developed later in the 19th century by
mathematicians including Venn~\cite{venn1866logic} and
Boole~\cite{boole1854investigation}, in reactions that may well have been
propelled by reasonable objections to flawed posterior-only methods.
Modern frequentist theory has a rigorous set-theoretic formulation in the
Kolmogorov axioms~\cite{
kolomogoroff1933de,
kolomogoroff1950translated,
axioms1010038
}.
Applied frequentist methods, unlike frequentist theory, are constrained with
considerably less precision --- the denial of certain (non-frequency)
probabilities does not construct their replacements, and various surrogates
have evolved.
Modern frequentist methods are largely attributable to the authors
R. A. Fisher (for likelihood, $p$-values, and estimation criteria), and
J. Neyman and E. S. Pearson
(for statistical hypothesis testing and confidence intervals).
Together these do not form a harmonious or consistent framework ---
the two groups have emphatic disagreements in print ---
but do cover many practices that continue to survive in modern science
and explain important features of the methods apply to LHC data, which will be
introduced in Section~\ref{sec:searches_practice}.

\TODO{Fisher's great contribution --- likelihood}
\TODO{
Fisher:
criteria for estimators,
(consistency, efficiency, sufficiency)
~\cite{fisher1922estimators}
testing, p-value,
likelihood, also ratios of p-values (here named probable errors)
(introduced the term likelihood? cite?)
~\cite{fisher1921probable}
tail area = significance
~\cite{fisher1925smrw}
}
Figure~\ref{fig:searches_fisher_likelihood}

@@
